---
title: "RR2 draft"
author: "Ouissal Akioui"
format: 
  html:
    self_contained: true
    toc: true
    toc-depth: 2
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

## PART 1: STM & REGRESSION

## Introduction

This report examines the thematic structure of cultural coverage in Moroccan digital news. The dataset consists of 15,966 articles drawn from four major online outlets — Hespress, Hibapress, Le360, and Akhbarona — filtered to the Culture category from a larger Arabic-language news corpus. Each observation corresponds to a single published article, with outlet identity recorded as the primary grouping variable.

Rather than treating cultural news as a homogeneous category, this analysis asks how cultural coverage is internally structured and whether outlets differ in their thematic priorities. I use a Structural Topic Model (STM) to identify 10 latent sub-themes within the corpus, ranging from cinema and music festivals to literary culture and institutional policy coverage. Topic proportions are then regressed on outlet identity to test the following hypotheses.

H1: Thematic emphasis in cultural news coverage differs significantly across Moroccan digital outlets.

H2: Le360 and Hespress will allocate greater coverage to institutional and literary cultural topics compared to Hibapress and Akhbarona, with Le360 leading in literary and intellectual themes given its formally oriented reputation, and Hespress following closely behind due to its perceived institutional character.

## Analyses

Four filtered datasets were merged into a single dataframe containing 15,966 articles and three variables: article title, body text, and source.

```{r}
rm(list=ls()) #This line of code cleans the environment

```

```{r}

library(tidyverse)

# Load all datasets with source labels
hespress   <- read_csv("Hespress.ma.csv",   show_col_types = FALSE) %>% mutate(source = "Hespress")
hibapress  <- read_csv("Hibapress.com.csv", show_col_types = FALSE) %>% mutate(source = "Hibapress")
le360      <- read_csv("Le360.com.csv",     show_col_types = FALSE) %>% mutate(source = "Le360")
akhbarona  <- read_csv("Akhbarona.ma.csv",  show_col_types = FALSE) %>% mutate(source = "Akhbarona")

```

```{r}
#filtering all of them to keep only culture
hespress_c  <- hespress  %>% filter(Category == "Culture")
hibapress_c <- hibapress %>% filter(Category == "Culture")
le360_c     <- le360     %>% filter(Category == "Culture")
akhbarona_c <- akhbarona %>% filter(Category == "Culture")


```

```{r}
#merging all of it in one with a newly created source variable
culture_df <- bind_rows(hespress_c, hibapress_c, le360_c, akhbarona_c)

table(culture_df$source)
nrow(culture_df)
```

```{r}
#remove category since there is just one
culture_df <- culture_df %>% select(-Category)
View(culture_df)

```

### SECTION 1 : TEXT PREPROCESSING FOR ARABIC TOPIC MODELLING

Before the topic model could be run, the raw article text required several cleaning steps to reduce noise and isolate meaningful content.

The first step was basic text normalization. URLs, email addresses, numbers, and punctuation were stripped from the text. Arabic diacritics (tashkeel, the small vowel marks written above and below letters) were also removed, as the same word can appear with or without them across articles and they would otherwise be treated as distinct tokens.

The text was then tokenized, split into individual words, producing one token per row, which is the format the topic model requires.

The most iterative part of preprocessing was stopword removal. Arabic presents a particular challenge here: the language has an enormous number of functional words, particles, and conjunctions that carry no thematic meaning, and standard stopword lists do not catch all of them. A base list from the `stopwords` package (stopwords-iso) was used as a starting point, but it required substantial manual extension. Words like "said", "this", "after", and common reporting verbs were added by hand after inspecting which high-frequency words were slipping through. The word "Morocco" (المغرب) was also removed, since every article in the corpus is about Morocco by definition, it appears so frequently that it would dominate topics without adding any discriminating information.

Finally, any token shorter than three characters was discarded, as these tend to be leftover particles or noise rather than meaningful words.

One decision worth noting: the text was **not stemmed**. In many languages, stemming (reducing words to their root form) is a useful normalization step. Arabic, however, is a root-and-pattern language where words are built from three-letter roots through internal vowel changes and affixes, not simple prefix or suffix removal. Standard stemming approaches designed for languages like English perform poorly on Arabic and can distort meaning. Keeping the full word forms was therefore a more conservative and reliable choice, at the cost of some vocabulary redundancy.

Remove punctuation (Arabic and Latin)

```         
text = str_remove_all(text, "[[:punct:]]"),
```

```{r}

library(tidyverse)
library(tidytext)
library(stopwords)
library(stringr)


# 1A: CLEAN THE RAW TEXT

culture_clean <- culture_df %>%
  mutate(
    text = Body,
    
    # Remove URLs
    text = str_remove_all(text, "https?://\\S+|www\\.\\S+"),
    
    # Remove email addresses
    text = str_remove_all(text, "\\S+@\\S+"),
    
    # Remove numbers
    text = str_remove_all(text, "[[:digit:]]"),
    
    # Remove Arabic diacritics (tashkeel - the small vowel marks)
    text = str_remove_all(text, "[\u064B-\u065F]"),
    
  ) %>%
  # Add a unique document ID — we'll need this later
  mutate(doc_id = row_number())

# Quick check
head(culture_clean$text, 3)

# 1B: TOKENIZE (one word per row)

culture_tokens <- culture_clean %>%
  select(doc_id, source, text) %>%
  unnest_tokens(
    output = word,   # new column name for each token
    input  = text,   # column to tokenize
    token  = "words" # split by words
  )

# How many tokens total?
nrow(culture_tokens)


# 1C: REMOVE STOPWORDS


# Get Arabic stopwords from the stopwords package
arabic_stopwords <- stopwords("ar", source = "stopwords-iso")

# Also add some common noise words manually
# (we expand this list as we see what slips through)

custom_stopwords <- c(
  "قال", "قالت", "أن", "التي", "الذي", "وقد", "كان", "كانت",
  "هذا", "هذه", "وهو", "وهي", "أو", "لا", "ما", "مع", "عن",
  "في", "من", "على", "إلى", "عند", "حيث", "بعد", "قبل",
  "هناك", "يوم", "خلال", "حول", "بين", "أكد", "أضاف",
  "انه","أنه", "مغرب", "عبر", "عبد"  # morocco "مغرب" is  is so frequent it's basically noise. 
)

custom_stopwords2 <- c("وذلك","أيضا","إلا","لأن", "أخرى","خاصة", "المغربية", "المغربي", "المغرب")

all_stopwords <- c(arabic_stopwords, custom_stopwords, custom_stopwords2)

# Filter them out
culture_tokens <- culture_tokens %>%
  filter(!word %in% all_stopwords)


# 1D: REMOVE SHORT TOKENS (less than 3 characters)
# These are usually noise — conjunctions, leftover particles

culture_tokens <- culture_tokens %>%
  filter(nchar(word) >= 3)



culture_tokens %>%
  count(word, sort = TRUE) %>%
  head(30)

```

After preprocessing, the corpus contains **4.4 million tokens** across the 15,966 articles. The 30 most frequent words are overwhelmingly thematic rather than noise, which means the cleaning pipeline worked well.

```{r}
# Sample 100 rows first THEN apply function — much faster
set.seed(123)
test_tokens <- culture_tokens[sample(nrow(culture_tokens), 100), ]

# Apply stemmer
test_tokens$word_stemmed <- my_arabic_stemmer(test_tokens$word)

# View result
test_tokens[, c("word", "word_stemmed")]
```

Before running the topic model, we save a small reference table that links each article to its outlet. This is needed later to tell the model which newspaper published each piece.

```{r}
#SAVE THE METADATA (we'll need source labels later for STM)

culture_meta <- culture_clean %>%
  select(doc_id, source)

```

### SECTION 2: THE STRUCTURAL TOPIC MODEL

A Structural Topic Model (STM) is a method that automatically discovers recurring themes in a large collection of texts. The basic idea is simple: words that tend to appear together probably belong to the same theme. Feed the model 15,966 articles and it will find, on its own, that "film", "director", and "screening" keep showing up in the same articles, and label that cluster a topic. You then inspect the top words and give the topic a human-readable name.

What makes STM different from simpler versions of this approach is that it lets you bring in extra information about each document, in this case, which outlet published it. This means the model can go beyond just finding themes and actually test whether outlet identity predicts how much of each theme an article contains. That's the bridge between the topic modeling and the hypothesis.

Before the topic model can run, the cleaned text needs to be converted into a numerical format the algorithm can process. This takes the form of a Document-Feature Matrix (DFM); a large table where each row represents one article, each column represents a word from the full vocabulary, and each cell records how many times that word appeared in that article. Most cells are zero, since any given article only uses a small fraction of the total vocabulary. This matrix is built directly from the cleaned tokens and then converted into the format the STM package expects.

```{r}

library(quanteda)

# Build dfm directly from tokens 
culture_dfm <- culture_tokens %>%
  count(doc_id, word) %>%
  cast_dfm(document = doc_id, term = word, value = n)

# Now convert to STM format — this will work
culture_stm <- convert(culture_dfm, to = "stm")

# Align metadata
meta_aligned <- culture_meta %>%
  filter(doc_id %in% as.integer(culture_stm$documents %>% names())) %>%
  arrange(doc_id)

```

The model is set to find 10 topics. An earlier exploratory run with 15 produced too much overlap between adjacent themes, cinema and film festivals kept bleeding into each other, for instance, so 10 is a cleaner, more interpretable solution.

```{r}
#| results: hide
#| message: false

library(stm)

# RUN K=15 (just for comparison figure)
if (!file.exists("stm_model_k15.rds")) {
  stm_model_k15 <- stm(
    documents  = culture_stm$documents,
    vocab      = culture_stm$vocab,
    K          = 15,
    prevalence = ~ source,
    data       = meta_aligned,
    max.em.its = 75,
    init.type  = "Spectral",
    verbose    = TRUE
  )
  saveRDS(stm_model_k15, "stm_model_k15.rds")
} else {
  stm_model_k15 <- readRDS("stm_model_k15.rds")
}

```

```{r}
# Topic proportions summary plot
plot(stm_model_k15, type = "summary", xlim = c(0, 0.3))
```

```{r}
#| results: hide
#| message: false

library(stm)

# Run STM
if (!file.exists("stm_model_k10.rds")) {
  stm_model_k10 <- stm(
    documents  = culture_stm$documents,
    vocab      = culture_stm$vocab,
    K          = 10,
    prevalence = ~ source,
    data       = meta_aligned,
    max.em.its = 75,
    init.type  = "Spectral",
    verbose    = TRUE
  )
  saveRDS(stm_model_k10, "stm_model_k10.rds")
} else {
  stm_model_k10 <- readRDS("stm_model_k10.rds")
}

# Make sure stm_model still points to k10 for rest of analysis
stm_model <- stm_model_k10

```

Following model convergence, the top words per topic were inspected to assign interpretive labels, and the expected topic proportions were examined across the full corpus before moving to outlet-level comparisons.

```{r}
#| eval: false
# Top words per topic
labelTopics(stm_model, n = 10)

```

```{r}

## Assign labels based on inspection
topic_labels <- c(
  "Cinema & Film Awards",        # Topic 1
  "Celebrity Scandals & Legal",  # Topic 2
  "Music & Songs",               # Topic 3
  "TV & Ramadan Series",         # Topic 4
  "Cultural Festivals & Theatre",# Topic 5
  "Books, Heritage & Museums",   # Topic 6
  "Poetry, Obituaries & Arts",   # Topic 7
  "Celebrity Social Media",      # Topic 8
  "Literary Cinema & Criticism", # Topic 9
  "Cultural Policy & Ministry"   # Topic 10
)
```

```{r}

# Topic proportions summary plot
plot(stm_model, type = "summary", xlim = c(0, 0.3))
```

***Figure 1**. Expected proportion of each topic across the 15,966-article corpus. Topics are ordered by prevalence. Labels reflect the three highest-probability words per topic.*

Music & Songs (Topic 3) dominates the corpus, accounting for roughly 20% of all cultural coverage, followed closely by Cultural Festivals & Theatre (Topic 5) and Books, Heritage & Museums (Topic 6). Entertainment-adjacent topics — celebrity scandals (Topic 2) and social media content (Topic 8) — occupy the middle range, while TV & Ramadan Series (Topic 4), Cultural Policy (Topic 10), and Literary Cinema (Topic 9) are the least represented themes overall. This distribution suggests that Moroccan cultural news is driven primarily by music and live events coverage, with institutional and policy-oriented content occupying a comparatively smaller share of the space.

Having confirmed that the topic model converged, the estimated topic proportions are extracted from the model object. Each article is represented as a vector of 10 values summing to one (10 topics), where each value reflects the share of that article's content attributed to a given topic. Outlet identity is then re-attached to this dataframe, producing the dataset used for all subsequent outlet-level comparisons.

```{r}
#| results: hide
# Extract topic proportions per document
topic_proportions <- as.data.frame(stm_model$theta)

# theta is the matrix where each row is an article 
# and each column is a topic — values are the proportions
colnames(topic_proportions) <- topic_labels

# Add source back
topic_proportions$source <- meta_aligned$source

# Peek at it
head(topic_proportions)
```

### SECTION 3: ESTIMATE EFFECT OF SOURCE ON TOPIC PREVALENCE

To examine whether thematic emphasis in cultural coverage varies across news outlets, I use the `estimateEffect` function from the STM package, which regresses estimated topic proportions on outlet identity. For each of the 10 topics, the model estimates how much each outlet deviates from the baseline in its coverage of that topic. Unlike a standard OLS regression applied to raw topic proportions, `estimateEffect` incorporates the posterior uncertainty from the topic model itself when computing standard errors, producing more honest estimates. The `uncertainty = "Global"` option propagates this uncertainty globally across the model. Source (news outlet) is the sole predictor, with Akhbarona serving as the reference category by default due to alphabetical ordering.

```{r}
#| results: hide
effect <- estimateEffect(
  formula     = 1:10 ~ source,
  stmobj      = stm_model,
  metadata    = meta_aligned,
  uncertainty = "Global"
)

# Summary
summary(effect, topics = 1:10)
```

```{r}
library(knitr)
library(kableExtra)

# Extract coefficients into a clean dataframe
reg_table <- data.frame()

for (t in 1:10) {
  coefs <- summary(effect, topics = t)$tables[[1]]
  rows <- coefs[c("sourceHespress", "sourceHibapress", "sourceLe360"), ]
  
  reg_table <- rbind(reg_table, data.frame(
    Topic    = topic_labels[t],
    Outlet   = gsub("source", "", rownames(rows)),
    Estimate = round(rows[, "Estimate"], 3),
    SE       = round(rows[, "Std. Error"], 3),
    p        = round(rows[, "Pr(>|t|)"], 3)
  ))
}

# Add significance stars
reg_table$p <- ifelse(reg_table$p < 0.001, "<.001 ***",
               ifelse(reg_table$p < 0.01,  "<.01 **",
               ifelse(reg_table$p < 0.05,  "<.05 *",
               as.character(reg_table$p))))

kable(reg_table, row.names = FALSE,
      col.names = c("Topic", "Outlet", "β", "SE", "p")) %>%
  kable_styling(full_width = FALSE, font_size = 12) %>%
  collapse_rows(columns = 1, valign = "top")
```

***Table 1.** presents the estimated coefficients, standard errors, and significance levels for all ten topics across the three remaining outlets.*

The regression results reveal substantial and statistically significant differences in thematic emphasis across all four outlets, with Akhbarona serving as the reference category. These results support H1: outlet identity is a meaningful predictor of cultural topic prevalence.

Turning to H2, the findings only partially confirm the hypothesis. As expected, Hespress does not behave like a typical mass-audience outlet. It allocates significantly more coverage to Books, Heritage & Museums (β = 0.088, p \< .001), Literary Cinema & Criticism (β = 0.074, p \< .001), and Poetry, Obituaries & Arts (β = 0.081, p \< .001) than Akhbarona — and crucially, more than Le360 on the literary and intellectual dimensions. Le360, which was expected to lead on serious cultural content, instead shows its strongest advantage on Cinema & Film Awards (β = 0.061, p \< .001), Cultural Festivals & Theatre (β = 0.123, p \< .001), and Music & Songs (β = 0.138, p \< .001) — a more event-driven, institutionalized profile rather than a literary one.

Hibapress presents the sharpest contrast to both. Its dominant signal is Celebrity Social Media (β = 0.177, p \< .001), while it significantly undercovers TV & Ramadan Series (β = -0.022, p \< .01) and Cultural Policy (β = -0.020, p \< .01). It is clearly the entertainment-oriented outlet.

The pattern that emerges is therefore more nuanced than a simple serious-versus-popular divide. Hespress occupies an unexpected literary niche, Le360 owns the institutional events space, and Hibapress corners celebrity content.

### SECTION 4: BOOTSTRAPPING

What we're doing is bootstrapping the regression step conditional on the topic model being fixed, instead of a bootstrap re-estimating the entire topic model on each bootstrap sample.

The regression estimates above rely on standard errors that incorporate posterior uncertainty from the topic model itself, but a further robustness check is warranted. To confirm that the outlet differences identified are stable and not an artifact of the specific sample, I complement the regression with a bootstrap procedure. Rather than re-estimating the full topic model on each iteration (which would be computationally a lot), the bootstrap resamples articles with replacement and refits the regression on each sample, holding the topic model fixed. This produces empirical confidence intervals that do not depend on distributional assumptions. If the bootstrapped intervals align with the regression results and do not straddle zero for the same coefficients, the findings can be considered robust.

```{r}

#| results: hide
#| message: false

set.seed(123)
n_boot <- 1000  # number of bootstrap iterations

# Store bootstrapped coefficients for each topic
boot_results <- list()

for (t in 1:10) {
  topic_prop <- stm_model$theta[, t]  # Extract topic proportion for this topic
  boot_data <- data.frame(        # Combine with source
    proportion = topic_prop,
    source     = meta_aligned$source
  )
   
  # Bootstrap
  boot_coefs <- replicate(n_boot, {
    idx    <- sample(nrow(boot_data), replace = TRUE) # Resample with replacement
    sample <- boot_data[idx, ]
    fit    <- lm(proportion ~ source, data = sample)  # Refit regression
    coef(fit)
  })
  boot_results[[t]] <- t(boot_coefs)
}

# Build summary table
boot_summary <- data.frame()

for (t in 1:10) {
  ci <- apply(boot_results[[t]], 2, quantile, probs = c(0.025, 0.975))
  for (src in c("sourceHespress", "sourceHibapress", "sourceLe360")) {
    boot_summary <- rbind(boot_summary, data.frame(
      Topic  = topic_labels[t],
      Outlet = gsub("source", "", src),
      Lower  = round(ci[1, src], 3),
      Upper  = round(ci[2, src], 3)
    ))
  }
}

```

```{r}
#| echo: false
kable(boot_summary, row.names = FALSE,
      col.names = c("Topic", "Outlet", "2.5%", "97.5%"),
      caption = "Table 2. Bootstrapped 95% Confidence Intervals (ref: Akhbarona)") %>%
  kable_styling(full_width = FALSE, font_size = 12) %>%
  collapse_rows(columns = 1, valign = "top")
```

The bootstrapped confidence intervals corroborate the `estimateEffect` results. Across all 10 topics, coefficients that were significant in the regression maintain the same sign and do not straddle zero in the bootstrap, confirming that the outlet differences identified are stable and not a sampling artifact. The most notable results directly relevant to H2 hold firmly: Hespress shows consistently positive intervals for Books, Heritage & Museums and Literary Cinema & Criticism, while Le360's intervals for the same topics are narrow and cross zero, confirming that Hespress genuinely leads on literary content.

```{r}
library(ggplot2)

# Build a dataframe of bootstrapped CIs
plot_data <- data.frame()

for (t in 1:10) {
  ci <- apply(boot_results[[t]], 2, quantile, probs = c(0.025, 0.5, 0.975))
  
  for (src in c("sourceHespress", "sourceHibapress", "sourceLe360")) {
    plot_data <- rbind(plot_data, data.frame(
      topic  = topic_labels[t],
      source = gsub("source", "", src),
      lo     = ci[1, src],
      mid    = ci[2, src],
      hi     = ci[3, src]
    ))
  }
}

# Plot
ggplot(plot_data, aes(x = mid, y = topic, color = source)) +
  geom_point(position = position_dodge(0.5), size = 2) +
  geom_errorbarh(aes(xmin = lo, xmax = hi),
                 position = position_dodge(0.5), height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ source) +
  labs(
    title = "Effect of News Outlet on Cultural Topic Prevalence",
    subtitle = "Bootstrapped 95% Confidence Intervals (ref: Akhbarona)",
    x = "Coefficient (difference in topic proportion vs Akhbarona)",
    y = NULL
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

***Figure 2.** Effect of news outlet on cultural topic prevalence, relative to Akhbarona (reference category). Points represent bootstrapped median coefficients and horizontal lines denote 95% confidence intervals. Coefficients to the right of zero indicate greater topic prevalence relative to Akhbarona; coefficients to the left indicate lower prevalence.*

Figure 2 visualises the bootstrapped regression results across all ten topics and three outlets. The pattern is striking in its consistency. Hespress (red) shows positive and precisely estimated effects for Books, Heritage & Museums, Literary Cinema & Criticism, and Poetry, Obituaries & Arts — its literary profile is unambiguous. Le360 (blue) dominates on Music & Songs, Cultural Festivals & Theatre, and Cinema & Film Awards, confirming its event-driven institutional orientation. Hibapress (green) stands out almost exclusively on Celebrity Social Media, where its coefficient dwarfs all others in the model, while remaining near zero or negative across every other topic.

These results broadly support H1: outlet identity does predict thematic emphasis, and the differences are statistically robust across nearly all topics.

The most direct test of H2 is visible in the Books, Heritage & Museums and Literary Cinema & Criticism rows: Hespress sits clearly to the right while Le360 hovers near zero, confirming that Hespress, not Le360, is the literary outlet in the Moroccan cultural news landscape according to this corpus. This is the core finding of the analysis.

## PART 2: PREDICTIVE MODELLING

**Research Question:** What article-level features predict the degree of celebrity focus in Moroccan cultural news articles?

Having established that outlet identity shapes thematic emphasis in Moroccan cultural news, I now extend the analysis into a predictive framework. The outcome variable is a composite celebrity focus score, constructed by summing each article's STM-estimated proportions on Topic 2 (Celebrity Scandals & Legal) and Topic 8 (Celebrity Social Media). These two topics were merged because they both capture celebrity-oriented content and treating them separately would artificially split the same underlying dimension.

Predictors include outlet identity, article length, title length, vocabulary richness, and the proportion of foreign-script characters (entre autres) in the body text. I first evaluate a baseline linear regression on a held-out test set, then use 10-fold cross-validation to obtain a more reliable estimate of predictive performance, and finally apply lasso regression to determine which subset of features produces the most predictive model.

### SECTION 1: BUILDING OUR PREDICTORS AND OUTCOME

Article length

```{r}
culture_df <- culture_df %>%
  mutate(article_length = str_count(Body, "\\S+"))
```

Title length

```{r}
culture_df <- culture_df %>%
  mutate(title_length = str_count(Title, "\\S+"))
```

Vocabulary richness

```{r}
culture_df <- culture_df %>% mutate(doc_id = row_number())
vocab_richness <- culture_tokens %>%
  group_by(doc_id) %>%
  summarise(
    total_words  = n(),
    unique_words = n_distinct(word),
    vocab_rich   = unique_words / total_words
  )
culture_df <- culture_df %>% left_join(vocab_richness, by = "doc_id")
```

Proportion of foreign words

```{r}
culture_df <- culture_df %>%
  mutate(
    latin_chars = str_count(Body, "[A-Za-z]"),
    total_chars = nchar(Body),
    foreign_prop = latin_chars / total_chars
  )
```

Topic proportions

```{r}
topic_props <- as.data.frame(stm_model$theta)
colnames(topic_props) <- paste0("topic_", 1:10)

meta_aligned <- bind_cols(meta_aligned, topic_props)

culture_df <- culture_df %>%
  left_join(
    meta_aligned %>% select(doc_id, starts_with("topic_")),
    by = "doc_id"
  )
```

Celebrity score: the outcome

```{r}

# Add celebrity_score as a new column in culture_df
culture_df <- culture_df %>%
  mutate(celebrity_score = topic_8 + topic_2)


```

Other text features

```{r}

# Add quick text features to culture_df
culture_df <- culture_df %>%
  mutate(
    
    # Average word length in characters — celebrity writing tends to be simpler
    avg_word_length = nchar(gsub("[^a-zA-Z\u0600-\u06FF]", "", Body)) / article_length,
    
    # Quote density — how often reporting verbs appear per word
    quote_density = str_count(Body, "قال|قالت|أكد|أكدت|صرح|أوضح") / article_length
  )
```

**Has number in title** — celebrity articles often have listicle-style titles ("5 things about...")

```{r}
culture_df <- culture_df %>%
  mutate(title_has_number = as.integer(str_detect(Title, "[0-9]")))
```

For the next set of variables created, I ask the question of what makes an article celebrity focused theoritically.

-   It's about a specific **person** (not an institution or event)

-   It has an **emotional or dramatic angle** (scandal, death, romance)

-   It's **reactive** (responding to something a celebrity did or said)

So meaningful dummy variables could be the following

```{r}

#Article mentions death/obituary — could capture the poetry/obituary topic bleeding into celebrity
culture_df <- culture_df %>%
  mutate(mentions_death = as.integer(str_detect(Body, "وفاة|توفي|رحيل|فقدنا")))


#Article mentions a prize or award — captures cinema awards topic
culture_df <- culture_df %>%
  mutate(mentions_award = as.integer(str_detect(Body, "جائزة|تكريم|فاز|حصل")))

#Article mentions social media platforms — directly relevant to celebrity social media topic
culture_df <- culture_df %>%
  mutate(mentions_socialmedia = as.integer(str_detect(Body, "يوتيوب|انستغرام|تيكتوك|فيسبوك|تويتر")))
```

As part of this analysis, I attempted to enrich the predictor set by extracting named entity information from the article texts using the CAMeL Tools Arabic NER model (ner-arabert), a BERT-based named entity recognition system trained on Modern Standard Arabic. The goal was to create two additional article-level features — person_count, measuring the raw number of person names mentioned per article, and person_density, normalizing that count by article length — on the premise that celebrity-focused articles systematically mention more named individuals than articles covering institutional or literary cultural topics. Due to computational constraints and processing time limitations on a dataset of 15,966 articles, the pipeline could not be completed within the timeframe of this submission. This approach will be pursued in the final version of the report, where it is expected to meaningfully improve the predictive performance of the model.

```{r}
# Load NER results from Python
ner_results <- read_csv("C:/Users/hp/Documents/DAT505/DAT505-RR2/ner_results.csv",
                        show_col_types = FALSE)

# Quick check — make sure it looks right
head(ner_results)
nrow(ner_results)

# Join to culture_df by doc_id
culture_df <- culture_df %>%
  left_join(ner_results, by = "doc_id")

# Verify the new columns are there
colnames(culture_df)

# Quick sanity check — distribution of person counts
summary(culture_df$person_count)
summary(culture_df$person_density)
```

The reason we need `model_df` separately is that `culture_df` still has lots of columns we don't want in the regression (Title, Body, doc_id, total_words, the other topic proportions etc.). `model_df` is a clean, lean version with only the outcome + predictors.

```{r}

  # Keep only the variables we need: outcome + predictors
 
model_df <- culture_df %>%
  select(celebrity_score, source, article_length, title_length,
         vocab_rich, foreign_prop, avg_word_length, quote_density) %>%
  na.omit()
```

### SECTION 2: Predictive Modelling of Celebrity Focus

#### Baseline Model Performance: Train/Test Split

The dataset is split into 80% for training and 20% for testing. A linear regression is fitted on the training set only, then used to predict celebrity scores on the test set (data the model has never seen). This tells us how well the model generalises beyond the data it was trained on.

```{r}
# TRAIN / TEST SPLIT 

set.seed(123) # for reproducibility

# Randomly select 80% of row indices for training
train_idx <- sample(1:nrow(model_df), size = 0.8 * nrow(model_df))

# Split into train and test sets
train <- model_df[train_idx, ]
test  <- model_df[-train_idx, ]
```

```{r}
# FIT MAIN MODEL ON TRAIN SET 

# Linear regression predicting celebrity_score from all predictors
main_model <- lm(celebrity_score ~ ., data = train)

# Predict on the test set
predictions <- predict(main_model, newdata = test)
```

```{r}
# EVALUATE PERFORMANCE

# RMSE: average prediction error (lower = better)
rmse <- sqrt(mean((test$celebrity_score - predictions)^2))

# R-squared: proportion of variance explained (higher = better)
r2 <- cor(test$celebrity_score, predictions)^2

# MAE: mean absolute error — average absolute difference between predicted and actual
mae <- mean(abs(test$celebrity_score - predictions))

# Print all metrics
cat("RMSE:", round(rmse, 4), "\n")
cat("R²:  ", round(r2, 4), "\n")
cat("MAE: ", round(mae, 4), "\n")

# Full model summary on train set — shows coefficients and significance
summary(main_model)
```

The model was trained on 80% of the data and evaluated on the remaining 20%. On the test set, it achieved :

-   RMSE of 0.2536 means predictions are off by about 0.25 units on average. Since celebrity_score ranges from 0 to 2, this is a moderate error.

<!-- -->

-   R² of 0.1565 means the model explains about 16% of variance in celebrity focus.

<!-- -->

-   MAE of 0.1897 confirms predictions deviate by roughly 0.19 units on average from the true score.

The model explains approximately 16% of variance in celebrity focus which is very small, but consistent with the noisy, high-dimensional nature of text-derived data. Turning to the coefficients estimated on the training set, all predictors reach statistical significance. Outlet identity emerges as the dominant predictor: all three outlets score lower on celebrity content than Akhbarona, with Le360 showing the largest negative coefficient (β = -0.25). Among text features, quote density shows the strongest positive association (β = 8.81), suggesting that articles quoting individuals more frequently tend to be more celebrity-focused. Average word length is negatively associated with celebrity score (β = -0.07), consistent with the expectation that celebrity writing uses simpler, shorter words. Foreign script proportion is also negative (β = -0.54), while vocabulary richness shows a small positive effect (β = 0.27).

#### CROSS VALIDATION

**CV with 10 FOLD**

A single train/test split depends on which particular 80% of the data ends up in training. This means a different random split could give slightly different results. 10-fold cross-validation addresses this by splitting the full dataset into 10 equal chunks, training on 9 and testing on 1, rotating through all 10 combinations and averaging the results. This gives a more reliable and stable estimate of predictive performance.

```{r}
library(caret)

# Define 10-fold cross-validation rules
train_control_cv <- trainControl(
  method = "cv",
  number = 10
)

# Launch model training with 10-fold CV
cv_model <- train(
  celebrity_score ~ .,
  data      = model_df,
  method    = "lm",
  metric    = "RMSE",
  trControl = train_control_cv
)

print(cv_model)
```

10-fold cross-validation was chosen given the large sample size (n = 15,965), which ensures each fold contains sufficient observations for stable estimation. Across the 10 folds, the model achieves an average RMSE of 0.247, R² of 0.171, and MAE of 0.186.

-   RMSE of 0.247 means that on average the model's predictions deviate from the true celebrity score by about 0.25 units.

-   MAE of 0.186 tells a similar story but is less sensitive to large errors: the typical prediction is off by roughly 0.19 units.

-   R² of 0.171 means the model accounts for about 17% of the variance in celebrity focus across articles.

    These results are closely consistent with the train/test estimates (RMSE = 0.254, R² = 0.157, MAE = 0.190). It confirms that the model's performance is stable across different data splits and is not driven by one lucky partition. The modest but consistent R² of around 17% suggests that outlet identity and structural text features carry genuine, if limited, predictive signal for celebrity focus.

**CV with BOOTSTRAP**

Bootstrap resampling repeatedly draws random samples with replacement from the full dataset, meaning some articles appear multiple times in a sample and others not at all. This is done 10 times, each time training on the bootstrap sample and testing on the original data. It's a different way of estimating model stability that doesn't require splitting the data into fixed folds.

```{r}
# Define bootstrap rules
train_control_boot <- trainControl(
  method = "boot",
  number = 10
)

# Launch model training with bootstrap
cv_model_boot <- train(
  celebrity_score ~ .,
  data      = model_df,
  method    = "lm",
  metric    = "RMSE",
  trControl = train_control_boot
)

print(cv_model_boot)
```

The bootstrap results are nearly identical to the 10-fold CV estimates: RMSE of 0.247, R² of 0.165, and MAE of 0.186. The fact that two fundamentally different resampling strategies — one splitting the data into fixed folds, the other drawing random samples with replacement — converge on the same performance figures is strong evidence that the model's predictive performance is genuine and stable. The consistency across train/test split, 10-fold CV, and bootstrap resampling all pointing to roughly RMSE ≈ 0.25 and R² ≈ 0.16 gives confidence that these estimates are not an artifact of any particular methodological choice.

#### LASSO REGRESSION FOR PREDICTION

Lasso regression works like linear regression but adds a penalty for complexity. It shrinks coefficients of less useful variables towards zero and can drop them entirely. This forces the model to identify the most predictive subset of variables rather than just using everything. The optimal penalty strength (lambda) is chosen via 10-fold CV, selecting the value that minimises prediction error.

```{r}
library(glmnet)

# Define CV rules for lasso
tc_lasso <- trainControl(
  method = "cv",
  number = 10
)

# Train lasso model — alpha=1 means lasso, lambda range searched via CV
model_lasso <- train(
  celebrity_score ~ .,
  data     = model_df,
  method   = "glmnet",
  tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 0.01, 0.0001)),
  metric   = "RMSE",
  trControl = tc_lasso
)

# Best lambda chosen by CV
model_lasso$bestTune

# Results at best lambda
model_lasso$results[which(model_lasso$results$lambda == model_lasso$finalModel$lambdaOpt), ]

# Which variables survived lasso?
coef(model_lasso$finalModel, model_lasso$finalModel$lambdaOpt)
```

The optimal lambda selected by cross-validation is 0.0001 which is a very small penalty indicating that lasso barely needed to shrink any coefficients. This is confirmed by the coefficients output: all nine predictors survived lasso with non-zero coefficients, meaning none were deemed redundant enough to drop. This tells us that every variable in the model carries at least some predictive signal, and that the model is not overfitted.

At this optimal lambda, the model achieves RMSE of 0.247, R² of 0.170, and MAE of 0.186, virtually identical to both the train/test split and cross-validation estimates. The stability of these figures across three different evaluation approaches is the central finding of this section: the model consistently explains around 17% of variance in celebrity focus regardless of how performance is assessed.

In terms of which variables matter most, the lasso coefficients mirror the linear regression findings. Outlet identity dominates: Le360 shows the largest negative coefficient (β = -0.25), followed by Hibapress (β = -0.19) and Hespress (β = -0.18), all relative to Akhbarona as the reference category. Among text features, quote density is by far the strongest predictor (β = 8.59) — articles that quote individuals more frequently are substantially more celebrity-focused. Foreign script proportion is negatively associated with celebrity content (β = -0.55), average word length shows a small negative effect (β = -0.07), and vocabulary richness a small positive one (β = 0.27). Article length and title length survive but with negligible coefficients, suggesting they carry minimal practical importance.

Overall, the lasso results confirm that the most predictive model available with these variables is essentially the full model. However, the ceiling on predictive performance at around R² = 0.17 points to the limits of structural text features alone, and motivates the planned incorporation of named entity (person) density in the final version of this report.
